---
title: 'Probability and Statistics: To p or not to p? Week 5 Assignment Two: Hypothesis
  Testing'
author: "Armulfo Perez"
date: "27/6/2020"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

You are to test the claim by a mineral water bottle manufacturer that its bottles contain an average of 1000 ml (1 litre). A random sample of n=12 bottles resulted in the measurements (in ml): 992, 1002, 1000, 1001, 998, 999, 1000, 995, 1003, 1001, 997 and 997.

It is assumed that the true variance of water in all bottles is $\sigma^2$ =1.5, and that the amount of water in bottles is normally distributed.

Test the manufacturer's claim at the 1% significance level (you may use Excel to calculate the ppp-value). Also, briefly comment on what the hypothesis test result means about the manufacturer's claim, and if an error might have occurred which type of error it would be.

In summary, the assignment requires:

1. The calculation of the sample mean from the raw observations
2. The formulation of the hypotheses, H0 and H1
3. Calculation of the test statistic value
4. Calculation of the p-value
5. A decision of whether or not to reject H0
6. An inferential conclusion about what the test result means
7. Indication of which type of error might have occurred.

## 1. The calculation of the sample mean from the raw observations
    

```{r cars}
sample_data = c(992, 1002, 1000, 1001, 998, 999, 1000, 995, 1003, 1001, 997, 997)
sample_mean = mean(sample_data)
sample_mean
```

## 2. The formulation of the hypotheses, H0 and H1

H0: The sample mean is equal to 1000ml to the 1% significance level. It is assumed that the true variance of water in all bottles is $\sigma^2$ =1.5, and that the amount of water in bottles is normally distributed.

H1: The sample mean is NOT equal to 1000ml to the 1% significance level.

## 3. Calculation of the test statistic value

The test statistics is the normalized mean of the sample, denoted by z

```{r }
mu = 1000
sigma = sqrt(1.5)
n = 12
z = (sample_mean-mu)/(sigma/sqrt(n))
z
```

## 4. Calculation of the p-value

For a two-sided hypotesis testing of means equaltiy, the p-value is twice the normal distribution value of the test statistic


```{r }
p_value = 2*pnorm(z)
p_value
```

## 5. A decision of whether or not to reject H0

The p-value, `r p_value`, is  significant at the 1% significance level, thus, I  reject the null hypothesis of means equllity.

## 6. An inferential conclusion about what the test result means

There is enough evidence to reject the hypothesis that the mean of the sample is equal to 1000ml at the  1% significance level.

## 7. Indication of which type of error might have occurred.

There is a chance of a type II error, and that the sample mean is equal to 1000ml.

