---
title: 'Assignment: R Markdown Presentation & Plotly'
author: "Armulfo Perez"
date: "21/5/2020"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(plotly)
library(tidyverse)
library(maps)
```

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

Plotly examples from [Interactive web-based data visualization with R, plotly, and shiny by 
Carson Sievert](https://plotly-r.com/index.html)

## Basic chart

```{r,fig.align='center', message = FALSE}
library(plotly)
fig <- plot_ly(midwest, x = ~percollege, color = ~state, type = "box")
ggplotly(fig)
```

## A simple line graph

```{r, fig.align='center', message = FALSE}
library(plotly)

df <- data.frame(x =  1:10, y = (1:10)^2)

p <- ggplot(df, aes(x = x, y = y)) + geom_line() + labs(x = "X", y = "Y", title = "y = x^2")

ggplotly(p)
```

## Mapbox

```{r, message = FALSE,warning=FALSE}
library(maps)
library(plotly)
plot_mapbox(maps::canada.cities) %>%
  add_markers(
    x = ~long, 
    y = ~lat, 
    size = ~pop, 
    color = ~country.etc,
    colors = "Accent",
    text = ~paste(name, pop),
    hoverinfo = "text"
  )
```

## OpenStreetMap Tiles

```{r fig.align='center', message = FALSE,warning=FALSE}
library(plotly)

us_cities = read.csv("https://raw.githubusercontent.com/plotly/datasets/master/us-cities-top-1k.csv")

fig <- us_cities 
fig <- fig %>%
  plot_ly(
    lat = ~lat,
    lon = ~lon,
    marker = list(color = "fuchsia"),
    type = 'scattermapbox',
    hovertext = us_cities[,"City"]) 
fig <- fig %>%
  layout(
    mapbox = list(
      style = 'open-street-map',
      zoom =2.5,
      center = list(lon = -88, lat = 34))) 

fig
```

## Base Map Below A Trace

```{r fig.align='center', message = FALSE,warning=FALSE}
library(plotly)

us_cities = read.csv("https://raw.githubusercontent.com/plotly/datasets/master/us-cities-top-1k.csv")

fig <- us_cities 
fig <- fig %>%
  plot_ly(
    lat = ~lat,
    lon = ~lon,
    type = "scattermapbox",
    hovertext = us_cities[,"City"],
    marker = list(color = "fuchsia")) 
fig <- fig %>%
  layout(mapbox= list(
    style = "white-bg",
    zoom = 3,
    center = list(lon = -93 ,lat= 41),
    layers = list(list(
      below = 'traces',
      sourcetype = "raster",
      source = list("https://basemap.nationalmap.gov/arcgis/rest/services/USGSImageryOnly/MapServer/tile/{z}/{y}/{x}")))))

fig
```


## Mapbox

```{r fig.align='center', message = FALSE,warning=FALSE}
map_data("world", "canada") %>%
   group_by(group) %>%
   plot_mapbox(x = ~long, y = ~lat) %>%
   add_polygons() %>%
   layout(
     mapbox = list(
       center = list(lat = ~median(lat), lon = ~median(long))
     )
   )
```

## Mapbox with access token

```{r fig.align='center', message = FALSE,warning=FALSE}
us_cities = read.csv("https://raw.githubusercontent.com/plotly/datasets/master/us-cities-top-1k.csv")

fig <- us_cities 
fig <- fig %>%
  plot_ly(
    lat = ~lat,
    lon = ~lon,
    marker = list(color = "fuchsia"),
    type = 'scattermapbox',
    hovertext = us_cities[,"City"]) 
fig <- fig %>%
  layout(
    mapbox = list(
      style = 'dark',
      zoom =2.5,
      center = list(lon = -88, lat = 34))) 
fig <- fig %>%
  config(mapboxAccessToken = Sys.getenv("MAPBOX_TOKEN"))

fig
```


## Histograms

```{r fig.align='center', message = FALSE,warning=FALSE}
p1 <- plot_ly(diamonds, x = ~price) %>%
  add_histogram(name = "plotly.js")

price_hist <- function(method = "FD") {
  h <- hist(diamonds$price, breaks = method, plot = FALSE)
  plot_ly(x = h$mids, y = h$counts) %>% add_bars(name = method)
}

subplot(
  p1, price_hist(), price_hist("Sturges"),  price_hist("Scott"),
  nrows = 4, shareX = TRUE
)
```


## Multiple numeric distributions

```{r fig.align='center', message = FALSE,warning=FALSE}
one_plot <- function(d) {
  plot_ly(d, x = ~price) %>%
    add_annotations(
      ~unique(clarity), x = 0.5, y = 1, 
      xref = "paper", yref = "paper", showarrow = FALSE
    )
}

diamonds %>%
  split(.$clarity) %>%
  lapply(one_plot) %>% 
  subplot(nrows = 2, shareX = TRUE, titleX = FALSE) %>%
  hide_legend()
```


## Boxplots

```{r fig.align='center', message = FALSE,warning=FALSE}
p <- plot_ly(diamonds, y = ~price, color = I("black"), 
             alpha = 0.1, boxpoints = "suspectedoutliers")
p1 <- p %>% add_boxplot(x = "Overall")
p2 <- p %>% add_boxplot(x = ~cut)
subplot(
  p1, p2, shareY = TRUE,
  widths = c(0.2, 0.8), margin = 0
) %>% hide_legend()
```

## Trellis display

```{r fig.align='center', message = FALSE,warning=FALSE}
plot_ly(diamonds, x = ~price, y = ~interaction(clarity, cut)) %>%
  add_boxplot(color = ~clarity) %>%
  layout(yaxis = list(title = ""))
```

## 2D analog

```{r fig.align='center', message = FALSE,warning=FALSE}
p <- plot_ly(diamonds, x = ~log(carat), y = ~log(price))
subplot(
  add_histogram2d(p) %>%
    colorbar(title = "default") %>%
    layout(xaxis = list(title = "default")),
  add_histogram2d(p, zsmooth = "best") %>%
    colorbar(title = "zsmooth") %>%
    layout(xaxis = list(title = "zsmooth")),
  add_histogram2d(p, nbinsx = 60, nbinsy = 60) %>%
    colorbar(title = "nbins") %>%
    layout(xaxis = list(title = "nbins")),
  shareY = TRUE, titleX = TRUE
)
```

## 3D surface

```{r fig.align='center', message = FALSE,warning=FALSE}
x <- seq_len(nrow(volcano)) + 100
y <- seq_len(ncol(volcano)) + 500
plot_ly() %>% add_surface(x = ~x, y = ~y, z = ~volcano)